\textbf{Advection versus diffusion in brain intraventricular transport}

### 2025-03-27 paper/main.tex
clean

\title{\textbf{Advection versus diffusion in brain intraventricular transport}}
%
%
% Author: Please give full first and last names for authors and
% include * after the name of all corresponding authors

@@

\begin{document}
\maketitle

% Abstract should be written in the present tense and impersonal style (i.e., avoid we), and be at most 200 words long
\begin{abstract}
Cerebrospinal fluid (CSF) is integral to brain function, as the fluid
provides mechanical support for the brain and the flow of CSF helps
distribute nutrients, neurotransmitters and metabolites throughout
the central nervous system. CSF flow is partly driven
by beating of motile cilia, hair-like structures that line the ependymal cell layer
of the walls of the brain ventricles. Despite the physiological importance of CSF flow,
the underlying mechanisms of cilia-mediated flow and transport
in the brain ventricles remain yet to be comprehensively resolved.
In this study, the role of motile cilia in brain intraventricular CSF flow and transport
is analyzed and evaluated. The geometry of the zebrafish larval brain ventricles,
for which detailed knowledge of cilia properties and CSF motion velocities
and frequencies exists, was used to develop finite element methods that model flow and transport.
The computational model is validated by in vivo experiments that monitor transport
of a photoconvertible protein secreted in the brain ventricles.
The results show that while cilia contribute to advection of large particles,
diffusion plays a significant role in the transport of small molecules such as proteins.
It is also demonstrated that cilia location and the geometry of the ventricular system impact solute distribution.
Altogether, this work presents a computational framework that can be applied to other
ventricular systems, together with new concepts of how molecules are transported within the brain and its ventricles.

@@

% Keywords: Please provide a minimum of three and a maximum of seven keywords, separated by commas
\noindent\small\hspace{.8725cm}\textbf{Keywords} cerebrospinal fluid, brain ventricles, cilia, transport, finite elements

@@

%%--------------RESULTS--------------%%
\section{Results}

@@

\subsection{Photoconversion data analysis}
The image data analysis was performed in MATLAB and the codes are
openly available~\fixme{\cite{zenodo-link-to-all-code}}.  The
photoconverted channel was aligned to correct for drift in $x,y$
directions using previously developed
codes~\cite{Reiten2017Motile-Cilia-MediatedComputations,
  Ringers2023NovelEpithelia}.  Only stable recordings were further
analyzed.  To increase the signal-to-noise ratio, the images were
downsampled by a block average using a resampling factor of 6. % (using the MATLAB function blockproc).
The change in fluorescence intensity
$\Delta F = (F-F_0)/F_0$ was
calculated for each pixel and each time point.  Here, the
value $F_0$ is the average fluorescence intensity for the baseline
(the first 10 time-lapse images) acquired before photoconversion.

@@

To obtain $\Delta F$ values for each region of interest (ROI), ROIs were
first drawn on the aligned and block-averaged time series (\Cref{fig:fig2}).
Six ROIs were drawn manually: ROI 1 around the location of photoconversion,
and ROIs 2--6 in different regions of the ventricular system.
The pixel values of the fluorescence intensity within one ROI were
averaged for each time point, and the relative change $\Delta F$ calculated with this averaged value.
To normalize the data with respect to photoconversion efficiency, the
$\Delta F$ curves for all ROIs were divided by the $\Delta F$ values
obtained for ROI 1 at the end of the experiment,
so that $\Delta F$ values for the photoconverted site ROI 1 ranged from
zero to one. Finally, we averaged the $\Delta F$ data from controls and from mutants.
%We report these mean values together with a one-standard deviation error band.

\subsection{Intraventricular injection of microbeads and particle tracking}
%To recover the flow fields shown in~\Crefpanels{fig:fig1}{b, c},

We also carried out
particle tracking in a larval zebrafish injected with fluorescent beads, as described in previous work~\cite{Olstad2019CiliaryDevelopment}.
Anesthetized 2 dpf zebrafish were injected with 1 nl of a mixture containing 0.1\% w/v fluorescent beads
(SPHERO Fluorescent Yellow Particles 1\% w/v, F = 0.16 mm) diluted
in 7.5 mg/ml 70 kDa rhodamine B isothiocyanate-dextran (RITCdextran; Sigma-Aldrich, R9379)
dissolved in artificial CSF. Artificial CSF composition was as follows:
124 mM NaCl, 22 mM D-(+)-Glucose, 2.0 mM KCl, 1.6 mM MgSO\textsubscript{4} $\cdot$ 7 H\textsubscript{2}O,
1.3 mM KH\textsubscript{2}PO\textsubscript{4}, 24 mM NaHCO\textsubscript{3},
2.0 mM CaCl\textsubscript{2} $\cdot$ 2 H\textsubscript{2}O.
The needles used for the injections were pulled with a
Sutter Instrument Co. Model P-2000 from thin-walled glass
capillaries (1.00 mm; VWR) and cut open with a forceps.
A volume of 1 nl of the solution was injected with a pressure injector
(Eppendorf Femtojet 4i) in the rostral rhombencephalic ventricle.
The pressure and time were calibrated for each needle using a 0.01 mm calibration slide.

@@

%% comparing these experimental velocity values
%% with the simulated, maximum velocity magnitude.

%% Both $\rr$ and the value $\tau=0.65$ mPa were chosen by 
%% calibration against experimentally measured flow speeds~\cite{Olstad2019CiliaryDevelopment}.

@@

\subsection{Estimation of diffusion coefficients via the Stokes-Einstein relation}
Motivated by an interest in molecules that are significant in neural development,
we study the transport of molecules resembling (i) the Dendra2 Fluorescent Protein (Dendra2)
used in the photoconversion, (ii) the Starmaker+Green Fluorescent Protein (STM-GFP) reported by
Jeong~\emph{et al.}~\cite{Jeong2024TheZebrafish}, and (iii) extracellular vesicles.
To estimate the diffusion coefficients of these, when not
available in the literature, we use the Stokes--Einstein relation
\begin{equation}
    D = \frac{k_B T}{6\pi \mu R},
    \label{eq:D_stokes_einstein}
\end{equation}
which is an expression for the diffusion coefficient of a spherical
molecule suspended in a viscous fluid~\cite{Einstein1905UberTeilchen}.
In~\Cref{eq:D_stokes_einstein}, $k_B = \num{1.38e-23}$ J/K is the Boltzmann constant,
$T = 310$ K is the absolute temperature, $\mu$ is the dynamic viscosity of the
CSF, and $R$ is the radius of the molecule.
For the extracellular vesicles, we calculate $D$ using~\Cref{eq:D_stokes_einstein}
with the relatively large $R=150$ nm~\cite{Moghassemi2024ExtracellularDecade}.
%Hence this radius covers the slowest diffusing exosomes,
%a choice made because smaller exosomes would have diffusion coefficients
%similar to the other molecules we consider.
For STM-GFP, we estimate the diffusion coefficient by extrapolating
the experimentally observed diffusion coefficient of
GFP~\cite{Swaminathan1997PhotobleachingDiffusion, Potma2001ReducedCells},
assuming that if $D$ scales with $1/R$ according to~\Cref{eq:D_stokes_einstein},
$D$ would scale with the cube root of the molecule mass~\cite{Goodhill1997DiffusionGuidance}.
The resulting diffusion coefficients
are $D_1$ (extracellular vesicles), $D_2$ (STM-GFP) and $D_3$ (Dendra2) as given by \Cref{tab:diff_coeff_table}.

%Implicit in this approach is the assumption that the proteins all have the same density.

@@

%-----DISCRETIZATION AND MODEL IMPLEMENTATION-----%
\subsection{Numerical approximation of the Stokes equations}
The Stokes equations (\Cref{eq:stokes_eqs}) are discretized with a finite element method
in space, while in time, we employ a first-order implicit Euler discretization~\cite{butcher2016numerical}.
We set the timestep size $\Delta t = 0.023$ s such that 20 timesteps represent one cardiac cycle.
For the spatial discretization, we use piecewise linear Brezzi-Douglas-Marini (BDM)
elements for the velocity.~\cite{Brezzi1985TwoProblems, Nedelec1986AIR3}
The BDM elements have degrees of freedom associated with integral moments of the
velocity component normal to the facets of elements,
thus we enforce the impermeability condition $\uu\cdot\nn=0$ strongly.
We impose the boundary conditions for both the tangential traction $\bsigpar$ and
the normal pressure $\Tilde{p}(t)$ weakly.
For the pressure, we use piecewise constant (zeroth order discontinuous Galerkin (DG)) elements.
The resulting total number of degrees of freedom at each time step is 1 009 874.
The element combination $\mathrm{BDM}_1$--$\mathrm{DG}_{0}$ yields a non-conforming
discretization scheme for the Stokes equations~\cite{Stenberg1989SomeEquations}.
Continuity in tangential velocities on interior facets
is weakly enforced with a penalty parameter $\gamma=10$~\cite{Hong2016AEquations}.
Since the divergence of the velocity space is a subspace of the pressure space for this scheme,
the divergence-free condition of the velocity is satisfied exactly, ensuring mass conservation.~\cite{Boffi2008FiniteProblem}

@@

%%------PHYSICAL EXPERIMENTS------%%
%\section{Experimental Methods}

@@

\medskip
\noindent \textbf{Data Availability Statement} \par
\noindent The data and implementations presented in this paper are licensed under \fixme{LICENSE}
and openly available at \fixme{REPO-LINK}.

\medskip
\noindent \textbf{Conflict of Interest Disclosure} \par
\noindent The authors declare no conflicts of interest.

%\medskip
%\textbf{Supporting Information} \par %Please delete the Suppporting Information statement
%if it is not applicable. Please supply Supporting Information in another file.
%Supporting information should not be provided in .tex format
%Supporting Information is available from the Wiley Online Library or from the author.

% References
%\medskip

% Use the following code if you wish to generate your bibliography with BibTeX;
% replace the string "MSP-template" below with the name(s) of
% the BibTeX data base(s) you want to use.
% The resulting bibliography-output (the content of the .bbl file)
% must be pasted back into this file before submission.
% Please also include your BibTeX data base file(s) in your submission
% so that we can re-run BibTeX if necessary.
%
\bibliographystyle{MSP}
\bibliography{ref}

>>>>

\title{\textbf{Advection versus diffusion in brain intraventricular transport}}

@@

\begin{document}

\maketitle

\begin{abstract}
Cerebrospinal fluid (CSF) is integral to brain function, as the fluid
provides mechanical support for the brain and the flow of CSF helps
distribute nutrients, neurotransmitters and metabolites throughout
the central nervous system. CSF flow is partly driven
by beating of motile cilia, hair-like structures that line the ependymal cell layer
of the walls of the brain ventricles. Despite the physiological importance of CSF flow,
the underlying mechanisms of cilia-mediated flow and transport
in the brain ventricles remain yet to be comprehensively resolved.
In this study, the role of motile cilia in brain intraventricular CSF flow and transport
is analyzed and evaluated. The geometry of the zebrafish larval brain ventricles,
for which detailed knowledge of cilia properties and CSF motion velocities
and frequencies exists, was used to develop finite element methods that model flow and transport.
The computational model is validated by in vivo experiments that monitor transport
of a photoconvertible protein secreted in the brain ventricles.
The results show that while cilia contribute to advection of large particles,
diffusion plays a significant role in the transport of small molecules such as proteins.
It is also demonstrated that cilia location and the geometry of the ventricular system impact solute distribution.
Altogether, this work presents a computational framework that can be applied to other
ventricular systems, together with new concepts of how molecules are transported within the brain and its ventricles.

@@

\noindent\small\hspace{.8725cm}\textbf{Keywords} cerebrospinal fluid, brain ventricles, cilia, transport, finite elements

@@

\section{Results}

@@

\subsection{Photoconversion data analysis}
The image data analysis was performed in MATLAB and the codes are
openly available~\fixme{\cite{zenodo-link-to-all-code}}.  The
photoconverted channel was aligned to correct for drift in $x,y$
directions using previously developed
codes~\cite{Reiten2017Motile-Cilia-MediatedComputations,
  Ringers2023NovelEpithelia}.  Only stable recordings were further
analyzed.  To increase the signal-to-noise ratio, the images were
downsampled by a block average using a resampling factor of 6. 
The change in fluorescence intensity
$\Delta F = (F-F_0)/F_0$ was
calculated for each pixel and each time point.  Here, the
value $F_0$ is the average fluorescence intensity for the baseline
(the first 10 time-lapse images) acquired before photoconversion.

@@

To obtain $\Delta F$ values for each region of interest (ROI), ROIs were
first drawn on the aligned and block-averaged time series (\Cref{fig:fig2}).
Six ROIs were drawn manually: ROI 1 around the location of photoconversion,
and ROIs 2--6 in different regions of the ventricular system.
The pixel values of the fluorescence intensity within one ROI were
averaged for each time point, and the relative change $\Delta F$ calculated with this averaged value.
To normalize the data with respect to photoconversion efficiency, the
$\Delta F$ curves for all ROIs were divided by the $\Delta F$ values
obtained for ROI 1 at the end of the experiment,
so that $\Delta F$ values for the photoconverted site ROI 1 ranged from
zero to one. Finally, we averaged the $\Delta F$ data from controls and from mutants.

\subsection{Intraventricular injection of microbeads and particle tracking}
We carried out particle tracking in a larval zebrafish injected with fluorescent beads,
as described in previous work~\cite{Olstad2019CiliaryDevelopment}.
Anesthetized 2 dpf zebrafish were injected with 1 nl of a mixture containing 0.1\% w/v fluorescent beads
(SPHERO Fluorescent Yellow Particles 1\% w/v, F = 0.16 mm) diluted
in 7.5 mg/ml 70 kDa rhodamine B isothiocyanate-dextran (RITCdextran; Sigma-Aldrich, R9379)
dissolved in artificial CSF. Artificial CSF composition was as follows:
124 mM NaCl, 22 mM D-(+)-Glucose, 2.0 mM KCl, 1.6 mM MgSO\textsubscript{4} $\cdot$ 7 H\textsubscript{2}O,
1.3 mM KH\textsubscript{2}PO\textsubscript{4}, 24 mM NaHCO\textsubscript{3},
2.0 mM CaCl\textsubscript{2} $\cdot$ 2 H\textsubscript{2}O.
The needles used for the injections were pulled with a
Sutter Instrument Co. Model P-2000 from thin-walled glass
capillaries (1.00 mm; VWR) and cut open with a forceps.
A volume of 1 nl of the solution was injected with a pressure injector
(Eppendorf Femtojet 4i) in the rostral rhombencephalic ventricle.
The pressure and time were calibrated for each needle using a 0.01 mm calibration slide.

@@

@@

\subsection{Estimation of diffusion coefficients via the Stokes-Einstein relation}
Motivated by an interest in molecules that are significant in neural development,
we study the transport of molecules resembling (i) the Dendra2 Fluorescent Protein (Dendra2)
used in the photoconversion, (ii) the Starmaker+Green Fluorescent Protein (STM-GFP) reported by
Jeong~\emph{et al.}~\cite{Jeong2024TheZebrafish}, and (iii) extracellular vesicles.
To estimate the diffusion coefficients of these, when not
available in the literature, we use the Stokes--Einstein relation
\begin{equation}
    D = \frac{k_B T}{6\pi \mu R},
    \label{eq:D_stokes_einstein}
\end{equation}
which is an expression for the diffusion coefficient of a spherical
molecule suspended in a viscous fluid~\cite{Einstein1905UberTeilchen}.
In~\Cref{eq:D_stokes_einstein}, $k_B = \num{1.38e-23}$ J/K is the Boltzmann constant,
$T = 310$ K is the absolute temperature, $\mu$ is the dynamic viscosity of the
CSF, and $R$ is the radius of the molecule.
For the extracellular vesicles, we calculate $D$ using~\Cref{eq:D_stokes_einstein}
with the relatively large $R=150$ nm~\cite{Moghassemi2024ExtracellularDecade}.
For STM-GFP, we estimate the diffusion coefficient by extrapolating
the experimentally observed diffusion coefficient of
GFP~\cite{Swaminathan1997PhotobleachingDiffusion, Potma2001ReducedCells},
assuming that if $D$ scales with $1/R$ according to~\Cref{eq:D_stokes_einstein},
$D$ would scale with the cube root of the molecule mass~\cite{Goodhill1997DiffusionGuidance}.
The resulting diffusion coefficients
are $D_1$ (extracellular vesicles), $D_2$ (STM-GFP) and $D_3$ (Dendra2) as given by \Cref{tab:diff_coeff_table}.

@@

\subsection{Numerical approximation of the Stokes equations}
The Stokes equations (\Cref{eq:stokes_eqs}) are discretized with a finite element method
in space, while in time, we employ a first-order implicit Euler discretization~\cite{butcher2016numerical}.
We set the timestep size $\Delta t = 0.023$ s such that 20 timesteps represent one cardiac cycle.
For the spatial discretization, we use piecewise linear Brezzi-Douglas-Marini (BDM)
elements for the velocity.~\cite{Brezzi1985TwoProblems, Nedelec1986AIR3}
The BDM elements have degrees of freedom associated with integral moments of the
velocity component normal to the facets of elements,
thus we enforce the impermeability condition $\uu\cdot\nn=0$ strongly.
We impose the boundary conditions for both the tangential traction $\bsigpar$ and
the normal pressure $\Tilde{p}(t)$ weakly.
For the pressure, we use piecewise constant (zeroth order discontinuous Galerkin (DG)) elements.
The resulting total number of degrees of freedom at each time step is 1 009 874.
The element combination $\mathrm{BDM}_1$--$\mathrm{DG}_{0}$ yields a non-conforming
discretization scheme for the Stokes equations~\cite{Stenberg1989SomeEquations}.
Continuity in tangential velocities on interior facets
is weakly enforced with a penalty parameter $\gamma=10$~\cite{Hong2016AEquations}.
Since the divergence of the velocity space is a subspace of the pressure space for this scheme,
the divergence-free condition of the velocity is satisfied exactly, ensuring mass conservation.~\cite{Boffi2008FiniteProblem}

@@

@@

% Data Availability Statement
\medskip
\noindent \textbf{Data Availability Statement} \par
\noindent The data and implementations presented in this paper are licensed under \fixme{LICENSE}
and openly available at \fixme{REPO-LINK}.

% Conflic of Interest Disclosure
\medskip
\noindent \textbf{Conflict of Interest Disclosure} \par
\noindent The authors declare no conflicts of interest.

% Bibliography
\bibliographystyle{MSP}
\bibliography{ref}
